%
% requirements.tex
%
% Copyright (C) 2020 by SpaceLab.
%
% FloripaSat-2 Documentation
%
% This work is licensed under the Creative Commons Attribution-ShareAlike 4.0
% International License. To view a copy of this license,
% visit http://creativecommons.org/licenses/by-sa/4.0/.
%

%
% \brief Mission requirements chapter.
%
% \author Gabriel Mariano Marcelino <gabriel.mm8@gmail.com>
%
% \institution Universidade Federal de Santa Catarina (UFSC)
%
% \version 0.1.0
%
% \date 2020/06/06
%

\chapter{Mission Requirements} \label{ch:requirements}

\begin{enumerate}
    \item The power system shall be able to harvest solar energy.
    \item The power system shall be able to store energy for use when FloripaSat-2 is eclipsed.
    \item The power system shall supply energy to all other modules.
    \item The data handling system shall communicate with the other modules and store their data.
    \item The communications system shall send a beacon signal periodically using VHF radio.
    \item The communications system shall send the CubeSat telemetry using UHF radio.
    \item The communications system shall be able to receive telecommands and respond to them accordingly.
    \item The attitude system shall be able to perform a 1-axis stabilization of the CubeSat.
    \item FloripaSat-2 shall have the capability to receive and execute a shutdown telecommand, therefore ceasing all transmissions.
    \item The downlink transmissions shall be done once at a time, either telemetry or beacon.
    \item The ground station shall operate under the proper radio frequency communication licenses.
    \item FLoripaSat-2 shall comply with international and Brazilian radio license agreements and restrictions.
    \item The team shall build and operate a ground station for full communication with FloripaSat-2.
\end{enumerate}
